---
title: "HW 3A"
output: html_document
---

```{r message = FALSE, warning = FALSE, error = FALSE}
#load libraries
library(rio)
library(ggplot2)
library(tidyverse)
library(naniar)
```

```{r}
worlddf <- import("world indicators exercise.csv") %>% 
  rename("Country_Name" = 1,
         "Series_code" = 4,
         "1960" = 5,
         "2018" = 63,
         "Series_name" = 3)
```

```{r}
# Remove any extra rows with invalid values (take a look at Country Name)
world <- worlddf[-c(67:69), ]

#Filter the data to just the United States and United Kingdom
worldusuk <- world %>% 
  filter(Country_Name == "United States" | Country_Name ==  "United Kingdom")

#drop the series code column
worldusuk <- worldusuk %>% 
  select(-c(Series_code))

#pivot all years of data into a single column
pivot_world <- worldusuk %>% 
  pivot_longer(cols = c("1960":"2018")) %>% 
  rename(year = name)

#this is me
#dropping the []
pivot_world$year <- gsub("\\[.*","", pivot_world$year)

#creating a unique indicator
pivot_world$ID <- seq.int(nrow(pivot_world))

#writing a csv so hopefully this could work in tableau??? it doenst???
#write.csv(pivot_world, "pivotedbutneedmore.csv")

#pivot wider to have indicator names as columns
pivot_part2 <- pivot_world %>% 
  pivot_wider(id_cols = ID, 
              names_from = Series_name,
              values_from = value)

#combining the pivot_part2 with the pivot_world
pivotall <- full_join(pivot_world, pivot_part2)

#transfering all .. to NA
pivotall2 <- replace_with_na_all(data = pivotall, condition = ~.x == "..")

#drop series name and id
pivotall2 <- pivotall2 %>% 
  select(-c(Series_name, ID, value))

# drop all but C02 emissions, population total, agricultural land
pivotall3 <- pivotall2 %>% 
  select(1, 2, 3, 6, 10, 12) %>% 
  rename(Agland = 4,
         co2 = 5, 
         pop = 6)

#have to make numeric
#rename stuff
# Create a derived column called CO2 emissions (kt per person) using the calculation: CO2 emissions(kt) / Population total
pivot_part2_co[, 1:3] <- sapply(pivot_part2_co[, 1:2], as.factor)

finaldataset <- pivot_part2_co %>% 
  as.numeric(Agland) %>% 
  as.numeric(co2) %>% 
  as.numeric(pop)
  mutate("C02 Emissions (kt per person)" = "CO2 emissions (kt)") 
```

Filter the data to just the United States and United Kingdom
Drop  the Series Code column
Pivot all years of data into a single column
Un-Pivot all indicator names from rows into columns
Drop all but the following indicators: CO2 emissions (kt), Population total,  Agricultural land (sq. km)
Create a derived column called CO2 emissions (kt per person) using the calculation: CO2 emissions(kt) / Population total